QOCO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2023-11-03 paper.tex
F_revision_6

\documentclass[10pt, conference,letterpaper]{IEEEtran}

@@

\author{
		Iman Rahmati, Hamed Shah-Mansouri, and Ali Movaghar
%		$^\dagger$Department of Computer Engineering, Sharif University of Technology, Tehran, Iran \\
%		$^\ddagger$Department of Electrical Engineering, Sharif University of Technology, Tehran, Iran \\
%		email: {\{iman.rahmati, hamedsh, movaghar\}}@sharif.edu
	}

@@

\begin{abstract}
In the realm of mobile edge computing (MEC), efficient computation task offloading plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand responsive and reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic and uncertain mobile environment. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a Markov decision process (MDP) to maximize the long-term QoE of each user individually. We propose a decentralized QoE-oriented computation offloading (QOCO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other devices. %We integrate advanced techniques, including Long Short-Term Memory (LSTM) and Dueling Double Deep Q-Network (DQN), to enhance the accuracy of QoE estimation in the algorithm. 
Through numerical studies, we evaluate the performance of QOCO. Simulation results validate that the QOCO algorithm efficiently exploits the computational resources of edge nodes. Consequently, it can complete 14\% more tasks and reduce task delay and energy consumption by 9\% and 6\%, respectively. These together contribute to a significant improvement of at least 37\% in average QoE compared to an existing algorithm.
\end{abstract}

@@

	\item \textit{DRL-based Offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption to quantify the MDs' QoE. We propose QOCO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. With a focus on the MD's battery level, our approach leverages deep Q-network (DQN)\cite{mnih2015human} and long short-term memory (LSTM) \cite{hochreiter1997long} to prioritize and strike an appropriate balance between QoE factors.

	\item \textit{Performance Evaluation:} We conduct comprehensive experiments to evaluate the QOCO's performance. The simulation results confirm that when compared with the potential game-based offloading algorithm (PGOA) \cite{yang2018distributed} and several benchmark methods, our algorithm effectively utilizes the processing capabilities of MDs and ENs, resulting in a substantial improvement in the average QoE of the system.
\end{itemize}

@@

@@

\subsubsection{Offloading Decision Algorithm at MD $i \in \mathcal{I}$}
 We analyze a series of episodes, where $N^{\text{ep}}$ denotes the number of them. At the beginning of each episode, if MD $i \in \mathcal{I}$ receives a new task $z_i(t)$, it initializes the state $\mathcal{S}_i(1)$ and sends an $\textit{UpdateRequest}$ to EN $j_i$. After receiving the requested vector $\theta_i^{\text{E}}$ of $\textit{Net}_i^{\text{E}}$ from EN $j_i$, MD $i$ chooses the following action for task $z_i(t)$.
\begin{alignat}{2}
	\hspace*{-2mm}\boldsymbol{a}_i(t) \hspace*{-0.5mm} = \hspace*{-0.5mm}
	\begin{cases} 
		\text{arg $\max_{\boldsymbol{a}\in \mathcal{A}}Q_i^{\text{E}}(\boldsymbol{s}_i(t), \boldsymbol{a}; \theta^{\text{E}}_i)$,} & \text{w.p. $1-\boldsymbol{\epsilon}$,} \\
 	\text{pick a random action from $\mathcal{A}$,} & \text{w.p. $\boldsymbol{\epsilon}$,}
	\end{cases}
	\label{26}  
\end{alignat}
where w.p. stands for with probability, and $\boldsymbol{\epsilon}$ represents the random exploration probability. The value of $Q_i^{\text{E}}(\boldsymbol{s}_i(t), \boldsymbol{a}; \theta^{\text{E}}_i)$ indicates the Q-value under the parameter $\theta^{\text{E}}_i$ of the neural network $\textit{Net}_i^{\text{E}}$. Specifically, the MD with a probability of $1 - \boldsymbol{\epsilon}$ selects the action associated with the highest Q-value under $\textit{Net}_i^{\text{E}}$ in the observed state $\boldsymbol{s}_i(t)$.
\begin{algorithm} [tbp]
	\caption{QOCO Algorithm (Offloading Decision)}\label{alg:cap}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\Require state space $\mathcal{S}$, action space $\mathcal{A}$
		\Ensure MD $i \in \mathcal{I}$ experience 
		\For {episode 1 to $N^{\text{ep}}$}
		\State Initialize $\boldsymbol{s}_i(1)$
		\For {time slot $t \in \mathcal{T}$}
		\If{MD $i$ receives a new task $z_i(t)$}
		\State Send an \textit{UpdateRequest} to EN $j_i$;
		\State Receive network parameter vector $\theta_i^{\text{E}}$;
		\State Select action $\boldsymbol{a}_i(t)$ based on \eqref{26};
		\EndIf
		\State Observe a set of QoEs $\{\boldsymbol{q}_i(t'), t' \in \mathcal{F}_i^t\}$;
		\State Observe\textcolor{white}{i}the next\textcolor{white}{i}state $\textbf{\textit{s}}_i(t+1)$;\textcolor{white}{i}
		\For {each task $z_i(t')$ where $t' \in \mathcal{F}_i^t$} 
		\State Send \hspace{-1mm} $(\boldsymbol{s}_i(t'), \boldsymbol{a}_i(t'), \boldsymbol{q}_i(t'), \boldsymbol{s}_i(t'\hspace{-1mm}+1))$ to EN $j_i$;
		\EndFor
		\EndFor
		\EndFor
	\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[tbp]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{222} 		
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{223}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.7cm}
	\caption{The overall energy consumption under different computation workloads: (a) task arrival rate; (b) the number of MDs.}
	\label{chart2}
\end{figure}
Figs.~\ref{chart2} (a) and ~\ref{chart2} (b) illustrate the overall energy consumption for different values of task arrival rate and the number of MDs, respectively. At the lower task arrival rate, the total energy consumption of all methods is close to each other. The total energy consumption increases when we have a higher task arrival rate.  
%When the task arrival rate increases, the total energy consumption of each method increases differently. It can be found that the production of more computational tasks leads to a continuous increase in energy consumption until the capacity is saturated, and the energy consumption remains constant at its maximum. The QOCO algorithm takes into account the battery level of the MD in its decision-making process. 
As can be observed from Fig.~\ref{chart2} (a), at task arrival rate 450, QOCO effectively reduces overall energy consumption by 18\% and 15\% compared to RD and PGOA, respectively, as it takes into account the battery level of the MD in its decision-making process. However, it consumes more energy compared to LC and FO because they do not utilize all computing resources. In particular, LC only uses the MD's computational resources, while FO utilizes the allocated EN computing resources.
 \begin{figure}[tbp]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{333} 		
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{334}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.7cm}
	\caption{The average delay under different computation workloads: (a) task arrival rate; (b) the number of MDs.}
	\label{chart3}
\end{figure} 
In Fig.~\ref{chart2} (b), an increasing trend in overall energy consumption is observed as the number of MDs increases since the number of resources available in the system increases, which leads to higher energy consumption. The QOCO algorithm consistently outperforms RD and PGOA methods in overall energy consumption, especially when there are a large number of MDs. Specifically, QOCO demonstrates a 27\% and 16\% reduction in overall energy consumption compared to RD and PGOA, respectively, when the number of MDs increases to 110.

@@

% Enfore empty line after bibliography
\ \\
%

\thanks{I. Rahmati and A. Movaghar are with the Department of Computer Engineering, Sharif University of Technology, Tehran, Iran (e-mail:$\{\text{iman.rahmati, movaghar} \}$@sharif.edu). 
}

\thanks{H. Shah-Mansouri is with the Department of Electrical Engineering, Sharif University of Technology, Tehran, Iran (email: hamedsh@sharif.edu)}

>>>>

\documentclass[10pt, journal,letterpaper]{IEEEtran}

@@

\author{\IEEEauthorblockN{
		Iman Rahmati, Hamed Shah-Mansouri, and Ali Movaghar\\}
%		$^\dagger$Department of Computer Engineering, Sharif University of Technology, Tehran, Iran \\
%		$^\ddagger$Department of Electrical Engineering, Sharif University of Technology, Tehran, Iran \\
%		email: {\{iman.rahmati, hamedsh, movaghar\}}@sharif.edu
\thanks{I. Rahmati and A. Movaghar are with the Department of Computer Engineering, Sharif University of Technology, Tehran, Iran (e-mail:$\{\text{iman.rahmati, movaghar} \}$@sharif.edu). 
}

\thanks{H. Shah-Mansouri is with the Department of Electrical Engineering, Sharif University of Technology, Tehran, Iran (email: hamedsh@sharif.edu)}
	}

@@

\begin{abstract}
In the realm of mobile edge computing (MEC), efficient computation task offloading plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand responsive and reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic and uncertain mobile environment. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a Markov decision process (MDP) to maximize the long-term QoE of each user individually. We propose a decentralized QoE-oriented computation offloading (QOCO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other devices. Through numerical studies, we evaluate the performance of QOCO. Simulation results validate that the QOCO algorithm efficiently exploits the computational resources of edge nodes. Consequently, it can complete 14\% more tasks and reduce task delay and energy consumption by 9\% and 6\%, respectively. These together contribute to a significant improvement of at least 37\% in average QoE compared to an existing algorithm.
\end{abstract}

@@

	\item \textit{DRL-based Offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption to quantify the MDs' QoE. We propose QOCO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. With a focus on the MD's battery level, our approach leverages deep Q-network (DQN) \cite{mnih2015human} and long short-term memory (LSTM) \cite{hochreiter1997long} to prioritize and strike an appropriate balance between QoE factors.

	\item \textit{Performance Evaluation:} We conduct comprehensive experiments to evaluate the QOCO's performance. The simulation results show that our algorithm effectively utilizes the processing capabilities of MDs and ENs, resulting in a substantial improvement in the average QoE of the system, when comparing with the potential game-based offloading algorithm (PGOA) \cite{yang2018distributed} and several benchmark methods.
\end{itemize}

@@

\begin{algorithm} [tbp]
	\caption{QOCO Algorithm (Offloading Decision)}\label{alg:cap}
	\begin{algorithmic}[1]
		\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
		\renewcommand{\algorithmicensure}{\textbf{Output:}}
		\Require state space $\mathcal{S}$, action space $\mathcal{A}$
		\Ensure MD $i \in \mathcal{I}$ experience 
		\For {episode 1 to $N^{\text{ep}}$}
		\State Initialize $\boldsymbol{s}_i(1)$
		\For {time slot $t \in \mathcal{T}$}
		\If{MD $i$ receives a new task $z_i(t)$}
		\State Send an \textit{UpdateRequest} to EN $j_i$;
		\State Receive network parameter vector $\theta_i^{\text{E}}$;
		\State Select action $\boldsymbol{a}_i(t)$ based on \eqref{26};
		\EndIf
		\State Observe a set of QoEs $\{\boldsymbol{q}_i(t'), t' \in \mathcal{F}_i^t\}$;
		\State Observe\textcolor{white}{i}the next\textcolor{white}{i}state $\textbf{\textit{s}}_i(t+1)$;\textcolor{white}{i}
		\For {each task $z_i(t')$ where $t' \in \mathcal{F}_i^t$} 
		\State Send \hspace{-1mm} $(\boldsymbol{s}_i(t'), \boldsymbol{a}_i(t'), \boldsymbol{q}_i(t'), \boldsymbol{s}_i(t'\hspace{-1mm}+1))$ to EN $j_i$;
		\EndFor
		\EndFor
		\EndFor
	\end{algorithmic}
\end{algorithm}

@@

\subsubsection{Offloading Decision Algorithm at MD $i \in \mathcal{I}$}
 We analyze a series of episodes, where $N^{\text{ep}}$ denotes the number of them. At the beginning of each episode, if MD $i \in \mathcal{I}$ receives a new task $z_i(t)$, it initializes the state $\mathcal{S}_i(1)$ and sends an $\textit{UpdateRequest}$ to EN $j_i$. After receiving the requested vector $\theta_i^{\text{E}}$ of $\textit{Net}_i^{\text{E}}$ from EN $j_i$, MD $i$ chooses the following action for task $z_i(t)$.
\begin{alignat}{2}
	\hspace*{-2mm}\boldsymbol{a}_i(t) \hspace*{-0.5mm} = \hspace*{-0.5mm}
	\begin{cases} 
		\text{arg $\max_{\boldsymbol{a}\in \mathcal{A}}Q_i^{\text{E}}(\boldsymbol{s}_i(t), \boldsymbol{a}; \theta^{\text{E}}_i)$,} & \text{w.p. $1-\boldsymbol{\epsilon}$,} \\
 	\text{pick a random action from $\mathcal{A}$,} & \text{w.p. $\boldsymbol{\epsilon}$,}
	\end{cases}
	\label{26}  
\end{alignat}
where w.p. stands for with probability, and $\boldsymbol{\epsilon}$ represents the random exploration probability. The value of $Q_i^{\text{E}}(\boldsymbol{s}_i(t), \boldsymbol{a}; \theta^{\text{E}}_i)$ indicates the Q-value under the parameter $\theta^{\text{E}}_i$ of the neural network $\textit{Net}_i^{\text{E}}$. Specifically, the MD with a probability of $1 - \boldsymbol{\epsilon}$ selects the action associated with the highest Q-value under $\textit{Net}_i^{\text{E}}$ in the observed state $\boldsymbol{s}_i(t)$.

@@

\begin{figure}[tbp]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{222} 		
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{223}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.7cm}
	\caption{The overall energy consumption under different computation workloads: (a) task arrival rate; (b) the number of MDs.}
	\label{chart2}
\end{figure}
Figs.~\ref{chart2} (a) and ~\ref{chart2} (b) illustrate the overall energy consumption for different values of task arrival rate and the number of MDs, respectively. At the lower task arrival rate, the total energy consumption of all methods is close to each other. The total energy consumption increases when we have a higher task arrival rate.  
%When the task arrival rate increases, the total energy consumption of each method increases differently. It can be found that the production of more computational tasks leads to a continuous increase in energy consumption until the capacity is saturated, and the energy consumption remains constant at its maximum. The QOCO algorithm takes into account the battery level of the MD in its decision-making process. 
As can be observed from Fig.~\ref{chart2} (a), at task arrival rate 450, QOCO effectively reduces overall energy consumption by 18\% and 15\% compared to RD and PGOA, respectively, as it takes into account the battery level of the MD in its decision-making process. However, it consumes more energy compared to LC and FO because they do not utilize all computing resources. In particular, LC only uses the MD's computational resources, while FO utilizes the allocated EN computing resources.
 \begin{figure}[t]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{333} 		
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{334}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.7cm}
	\caption{The average delay under different computation workloads: (a) task arrival rate; (b) the number of MDs.}
	\label{chart3}
\end{figure} 
In Fig.~\ref{chart2} (b), an increasing trend in overall energy consumption is observed as the number of MDs increases since the number of resources available in the system increases, which leads to higher energy consumption. The QOCO algorithm consistently outperforms RD and PGOA methods in overall energy consumption, especially when there are a large number of MDs. Specifically, QOCO demonstrates a 27\% and 16\% reduction in overall energy consumption compared to RD and PGOA, respectively, when the number of MDs increases to 110.

@@
